---
title: "Math 664 Project"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
library(readr)
library(ggplot2)
library(mosaic)
library(dplyr)
library(mosaicCore)
library(RColorBrewer)
HateCrimes <- read_csv("NYPD_Hate_Crimes.csv")
hate_crimes <- read_csv("NYPD_Hate_Crimes.csv")
colnames(hate_crimes) <- c("Full.Complaint.ID", "Complaint.Year.Number", "Month.Number", "Record.Create.Date", "Complaint.Precinct.Code", "Patrol.Borough.Name", "County", "Law.Code.Category.Description", "Offense.Description", "PD.Code.Description", "Bias.Motive.Description", "Arrest.Date", "Arrest.ID")
```

```{r}
#replace blank values with "NA"; Arrest Id and Arrest Date each have 860 NA values
hate_crimes[hate_crimes == ""] <- NA 
#--colSums(is.na(hate_crimes))--
#factor some columns
hate_crimes$Patrol.Borough.Name<-as.factor(hate_crimes$Patrol.Borough.Name)
#turn Arrest Date into Date class 
hate_crimes$Arrest.Date<-as.Date(hate_crimes$Arrest.Date,format="%m/%d/%Y")
#Add another column to signify if arrested or not
hate_crimes$ArrestStatus <- ifelse(is.na(hate_crimes$Arrest.Date), "No Arrest", "Arrested")

#rename levels of boros (Patrol Borough Name)
levels(hate_crimes$Patrol.Borough.Name)[1]<-"North Brooklyn"
levels(hate_crimes$Patrol.Borough.Name)[2]<-"South Brooklyn"
levels(hate_crimes$Patrol.Borough.Name)[3]<-"Bronx"
levels(hate_crimes$Patrol.Borough.Name)[4]<-"North Manhattan"
levels(hate_crimes$Patrol.Borough.Name)[5]<-"South Manhattan"
levels(hate_crimes$Patrol.Borough.Name)[6]<-"North Queens"
levels(hate_crimes$Patrol.Borough.Name)[7]<-"South Queens"
levels(hate_crimes$Patrol.Borough.Name)[8]<-"Staten Island"

```











## What Assaults are most common among each Bias/Borough Group?
### Bias
```{r, fig.align='center'}
bias <- HateCrimes$`Bias Motive Description`
bias <- as.data.frame(table(bias))
bias1 <- bias[order(bias$Freq),]
view(bias1)
```

```{r}
## Make the frequencies numbers (rather than factors)
bias1$Freq <- as.numeric(as.character(bias1$Freq))
## Find a range of y's that'll leave sufficient space above the tallest bar
ylim <- c(0, 1.1*max(bias1$Freq))
## Plot, and store x-coordinates of bars in xx
xx <- barplot(bias1$Freq, xaxt = 'n', xlab = '', width = 0.85, ylim = ylim,
              main = "Bar Chart of Bias Motives", 
              ylab = "Frequency",
              col = "cadetblue2")
## Add text at top of bars
text(x = xx, y = bias1$Freq, label = bias1$Freq, pos = 3, cex = 0.8, col = "black")
## Add x-axis labels 
axis(1, at=xx, labels=bias1$bias, tick=FALSE, las=2, line=-0.5, cex.axis=0.5)

```

```{r}
sixtyyrs <- HateCrimes %>% filter(`Bias Motive Description` == "60 YRS AND OLDER")
anti_arab <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-ARAB")
anti_asain <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-ASIAN")
anti_black <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-BLACK")
anti_buddhist <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-BUDDHIST")
anti_catholic <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-CATHOLIC")
anti_female <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-FEMALE")
anti_lesbian <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-FEMALE HOMOSEXUAL (LESBIAN)")
anti_gendernonconforming <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-GENDER NON-CONFORMING")
anti_hindu <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-HINDU")
anti_hispanic <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-HISPANIC")
anti_jehovas<- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-JEHOVAHS WITNESS")
anti_jewish <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-JEWISH")
anti_lgbt <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-LGBT (MIXED GROUP)")
anti_gay <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-MALE HOMOSEXUAL (GAY)")
anti_multiracial <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-MULTI-RACIAL GROUPS")
anti_muslim <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-MUSLIM")
anti_otherethnicity <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-OTHER ETHNICITY")
anti_otherreligion <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-OTHER RELIGION")
anti_physicaldis <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-PHYSICAL DISABILITY")
anti_religion <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-RELIGIOUS PRACTICE GENERALLY")
anti_trans <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-TRANSGENDER")
anti_white <- HateCrimes %>% filter(`Bias Motive Description` == "ANTI-WHITE")
```

#### 60 Yrs
```{r}
assaults_60yrs <- as.vector(sixtyyrs$`Offense Description`) 

assaults_60yrs <- as.matrix(table(assaults_60yrs))
assaults_60yrs
```

#### Anti-Arab
```{r}
assaults_anti_arab <- as.vector(anti_arab$`Offense Description`) 

assaults_anti_arab <- as.matrix(table(assaults_anti_arab))
assaults_anti_arab

barchart(assaults_anti_arab, main = "Bar Chart of Assaults with Anti-Arab Bias" , col = "cadetblue2")
```

#### Anti-Asain
```{r}
assaults_anti_asain <- as.vector(anti_asain$`Offense Description`) 

assaults_anti_asain<- as.matrix(table(assaults_anti_asain))
assaults_anti_asain

barchart(assaults_anti_asain, main = "Bar Chart of Assaults with Anti-Asain Bias" , col = "cadetblue2")
```

#### Anti-Black
```{r}
assaults_anti_black <- as.vector(anti_black$`Offense Description`) 

assaults_anti_black<- as.matrix(table(assaults_anti_black))
assaults_anti_black

barchart(assaults_anti_black, main = "Bar Chart of Assaults with Anti-Black Bias" , col = "cadetblue2")
```

#### Anti-Buddhist
```{r}
assaults_anti_buddhist <- as.vector(anti_buddhist$`Offense Description`) 

assaults_anti_buddhist<- as.matrix(table(assaults_anti_buddhist))
assaults_anti_buddhist

barchart(assaults_anti_buddhist, main = "Bar Chart of Assaults with Anti_Buddhist Bias" , col = "cadetblue2")
```

#### Anti-Catholic
```{r}
assaults_anti_catholic <- as.vector(anti_catholic$`Offense Description`) 

assaults_anti_catholic<- as.matrix(table(assaults_anti_catholic))
assaults_anti_catholic

barchart(assaults_anti_catholic, main = "Bar Chart of Assaults with Anti-Catholic Bias" , col = "cadetblue2")
```

#### Anti-Female
```{r}
assaults_anti_female <- as.vector(anti_female$`Offense Description`) 

assaults_anti_female<- as.matrix(table(assaults_anti_female))
assaults_anti_female

barchart(assaults_anti_female, main = "Bar Chart of Assaults with Anti-Female Bias" , col = "cadetblue2")
```

#### Anti-Lesbian
```{r}
assaults_anti_lesbian <- as.vector(anti_lesbian$`Offense Description`) 

assaults_anti_lesbian<- as.matrix(table(assaults_anti_lesbian))
assaults_anti_lesbian

barchart(assaults_anti_lesbian, main = "Bar Chart of Assaults with Anti-Lesbian Bias" , col = "cadetblue2")
```

#### Anti-Gender Non-Conforming
```{r}
assaults_anti_gendernonconforming <- as.vector(anti_gendernonconforming$`Offense Description`) 

assaults_anti_gendernonconforming<- as.matrix(table(assaults_anti_gendernonconforming))
assaults_anti_gendernonconforming

barchart(assaults_anti_gendernonconforming, main = "Bar Chart of Assaults with Anti-Gender Non-Conforming Bias" , col = "cadetblue2")
```

#### Anti-Hindu
```{r}
assaults_anti_hindu <- as.vector(anti_hindu$`Offense Description`) 

assaults_anti_hindu<- as.matrix(table(assaults_anti_hindu))
assaults_anti_hindu

barchart(assaults_anti_hindu, main = "Bar Chart of Assaults with Anti-Hindu Bias" , col = "cadetblue2")
```

#### Anti-Hispanic
```{r}
assaults_anti_hispanic <- as.vector(anti_hispanic$`Offense Description`) 

assaults_anti_hispanic<- as.matrix(table(assaults_anti_hispanic))
assaults_anti_hispanic

barchart(assaults_anti_hispanic, main = "Bar Chart of Assaults with Anti-Hispanic Bias" , col = "cadetblue2")
```

#### Anti-Jehovas Witnesses
```{r}
assaults_anti_jehovas <- as.vector(anti_jehovas$`Offense Description`) 

assaults_anti_jehovas<- as.matrix(table(assaults_anti_jehovas))
assaults_anti_jehovas

barchart(assaults_anti_jehovas, main = "Bar Chart of Assaults with Anti-Jehovas Witnesses Bias" , col = "cadetblue2")
```

#### Anti-Jewish
```{r}
assaults_anti_jewish <- as.vector(anti_jewish$`Offense Description`) 

assaults_anti_jewish<- as.matrix(table(assaults_anti_jewish))
assaults_anti_jewish

barchart(assaults_anti_jewish, main = "Bar Chart of Assaults with Anti-Jewish Bias" , col = "cadetblue2")
```

#### Anti-LGBT
```{r}
assaults_anti_lgbt <- as.vector(anti_lgbt$`Offense Description`) 

assaults_anti_lgbt<- as.matrix(table(assaults_anti_lgbt))
assaults_anti_lgbt

barchart(assaults_anti_lgbt, main = "Bar Chart of Assaults with Anti-LGBT Bias" , col = "cadetblue2")
```

#### Anti-Gay
```{r}
assaults_anti_gay <- as.vector(anti_gay$`Offense Description`) 

assaults_anti_gay<- as.matrix(table(assaults_anti_gay))
assaults_anti_gay

barchart(assaults_anti_gay, main = "Bar Chart of Assaults with Anti-Gay Bias" , col = "cadetblue2")
```

#### Anti-Multiracial Groups
```{r}
assaults_anti_multiracial <- as.vector(anti_multiracial$`Offense Description`) 

assaults_anti_multiracial<- as.matrix(table(assaults_anti_multiracial))
assaults_anti_multiracial

barchart(assaults_anti_multiracial, main = "Bar Chart of Assaults with Anti-Multiracial Groups Bias" , col = "cadetblue2")
```

#### Anti-Muslim
```{r}
assaults_anti_muslim <- as.vector(anti_muslim$`Offense Description`) 

assaults_anti_muslim<- as.matrix(table(assaults_anti_muslim))
assaults_anti_muslim

barchart(assaults_anti_muslim, main = "Bar Chart of Assaults with Anti-Muslim Bias" , col = "cadetblue2")
```

#### Anti-Other Ethnicity
```{r}
assaults_anti_otherethnicity <- as.vector(anti_otherethnicity$`Offense Description`) 

assaults_anti_otherethnicity<- as.matrix(table(assaults_anti_otherethnicity))
assaults_anti_otherethnicity

barchart(assaults_anti_otherethnicity, main = "Bar Chart of Assaults with Anti-Other Ethnicity Bias" , col = "cadetblue2")
```

#### Anti-Other Religion
```{r}
assaults_anti_otherreligion <- as.vector(anti_otherreligion$`Offense Description`) 

assaults_anti_otherreligion<- as.matrix(table(assaults_anti_otherreligion))
assaults_anti_otherreligion

barchart(assaults_anti_otherreligion, main = "Bar Chart of Assaults with Anti-Other Religion Bias" , col = "cadetblue2")
```

#### Anti-Physical Disability
```{r}
assaults_anti_physicaldis <- as.vector(anti_physicaldis$`Offense Description`) 

assaults_anti_physicaldis<- as.matrix(table(assaults_anti_physicaldis))
assaults_anti_physicaldis

barchart(assaults_anti_physicaldis, main = "Bar Chart of Assaults with Anti-Physical Disability Bias" , col = "cadetblue2")
```

#### Anti-Religion Practices Generally
```{r}
assaults_anti_religion <- as.vector(anti_religion$`Offense Description`) 

assaults_anti_religion<- as.matrix(table(assaults_anti_religion))
assaults_anti_religion

barchart(assaults_anti_religion, main = "Bar Chart of Assaults with Anti-Religion Practices Generally Bias" , col = "cadetblue2")
```

#### Anti-Transgender
```{r}
assaults_anti_trans <- as.vector(anti_trans$`Offense Description`) 

assaults_anti_trans<- as.matrix(table(assaults_anti_trans))
assaults_anti_trans

barchart(assaults_anti_trans, main = "Bar Chart of Assaults with Anti-Transgender Bias" , col = "cadetblue2")
```

#### Anti-White
```{r}
assaults_anti_white <- as.vector(anti_white$`Offense Description`) 

assaults_anti_white<- as.matrix(table(assaults_anti_white))
assaults_anti_white

barchart(assaults_anti_white, main = "Bar Chart of Assaults with Anti-White Bias" , col = "cadetblue2")
```

### Borough
```{r}
borough <- HateCrimes$`Patrol Borough Name`
borough <- as.data.frame(table(borough))
borough1 <- borough[order(borough$Freq),]
view(borough1)
```


```{r}
## Make the frequencies numbers (rather than factors)
borough1$Freq <- as.numeric(as.character(borough1$Freq))
## Find a range of y's that'll leave sufficient space above the tallest bar
ylim1 <- c(0, 1.1*max(borough1$Freq))
## Plot, and store x-coordinates of bars in xx
xx1 <- barplot(borough1$Freq, xaxt = 'n', xlab = '', width = 0.85, ylim = ylim1,
              main = "Bar Chart of Hate Crimes in Each Borough", 
              ylab = "Frequency",
              col = "cadetblue2")
## Add text at top of bars
text(x = xx1, y = borough1$Freq, label = borough1$Freq, pos = 3, cex = 0.8, col = "black")
## Add x-axis labels 
axis(1, at=xx1, labels=borough1$borough, tick=FALSE, las=2, line=-0.5, cex.axis=0.5)

```


```{r}
bklyn_north <- HateCrimes %>% filter(`Patrol Borough Name` == "PATROL BORO BKLYN NORTH")
bklyn_south <- HateCrimes %>% filter(`Patrol Borough Name` == "PATROL BORO BKLYN SOUTH")
bronx <- HateCrimes %>% filter(`Patrol Borough Name` == "PATROL BORO BRONX")
man_north <- HateCrimes %>% filter(`Patrol Borough Name` == "PATROL BORO MAN NORTH")
man_south <- HateCrimes %>% filter(`Patrol Borough Name` == "PATROL BORO MAN SOUTH")
queens_north <- HateCrimes %>% filter(`Patrol Borough Name` == "PATROL BORO QUEENS NORTH")
queens_south <- HateCrimes %>% filter(`Patrol Borough Name` == "PATROL BORO QUEENS SOUTH")
staten_island <- HateCrimes %>% filter(`Patrol Borough Name` == "PATROL BORO STATEN ISLAND")
```

#### Brooklyn North
```{r}
assaults_bklyn_north <- as.vector(bklyn_north$`Offense Description`) 

assaults_bklyn_north<- as.matrix(table(assaults_bklyn_north))
assaults_bklyn_north

barchart(assaults_bklyn_north, main = "Bar Chart of Assaults in Brooklyn North" , col = "cadetblue2")
```

#### Brooklyn South
```{r}
assaults_bklyn_south <- as.vector(bklyn_south$`Offense Description`) 

assaults_bklyn_south<- as.matrix(table(assaults_bklyn_south))
assaults_bklyn_south

barchart(assaults_bklyn_south, main = "Bar Chart of Assaults in Brooklyn South" , col = "cadetblue2")
```

#### Bronx
```{r}
assaults_bronx <- as.vector(bronx$`Offense Description`) 

assaults_bronx<- as.matrix(table(assaults_bronx))
assaults_bronx

barchart(assaults_bronx, main = "Bar Chart of Assaults in Bronx" , col = "cadetblue2")
```

#### Manhattan North
```{r}
assaults_man_north <- as.vector(man_north$`Offense Description`) 

assaults_man_north<- as.matrix(table(assaults_man_north))
assaults_man_north

barchart(assaults_man_north, main = "Bar Chart of Assaults in Manhattan North" , col = "cadetblue2")
```

#### Manhattan South
```{r}
assaults_man_south <- as.vector(man_south$`Offense Description`) 

assaults_man_south<- as.matrix(table(assaults_man_south))
assaults_man_south

barchart(assaults_man_south, main = "Bar Chart of Assaults in Manhattan South" , col = "cadetblue2")
```

#### Queens North
```{r}
assaults_queens_north <- as.vector(queens_north$`Offense Description`) 

assaults_queens_north<- as.matrix(table(assaults_queens_north))
assaults_queens_north

barchart(assaults_queens_north, main = "Bar Chart of Assaults in Queens North" , col = "cadetblue2")
```

#### Queens South
```{r}
assaults_queens_south <- as.vector(queens_south$`Offense Description`) 

assaults_queens_south<- as.matrix(table(assaults_queens_south))
assaults_queens_south

barchart(assaults_queens_south, main = "Bar Chart of Assaults in Queens South" , col = "cadetblue2")
```

#### Staten Island
```{r}
assaults_staten_island <- as.vector(staten_island$`Offense Description`) 

assaults_staten_island<- as.matrix(table(assaults_staten_island))
assaults_staten_island

barchart(assaults_staten_island, main = "Bar Chart of Assaults in Staten Island" , col = "cadetblue2")
```

### Bias group assaulted the most in each borough
#### Brooklyn North
```{r}
bias_bklyn_north <- as.vector(bklyn_north$`Bias Motive Description`) 

bias_bklyn_north<- as.matrix(table(bias_bklyn_north))
bias_bklyn_north

barchart(bias_bklyn_north, main = "Bar Chart of Bias in Brooklyn North" , col = "cadetblue2")
```

#### Brooklyn South
```{r}
bias_bklyn_south <- as.vector(bklyn_south$`Bias Motive Description`) 

bias_bklyn_south<- as.matrix(table(bias_bklyn_south))
bias_bklyn_south

barchart(bias_bklyn_south, main = "Bar Chart of Bias in Brooklyn South" , col = "cadetblue2")
```

#### Bronx
```{r}
bias_bronx <- as.vector(bronx$`Bias Motive Description`) 

bias_bronx<- as.matrix(table(bias_bronx))
bias_bronx

barchart(bias_bronx, main = "Bar Chart of Bias in Bronx" , col = "cadetblue2")
```

#### Manhattan North
```{r}
bias_man_north <- as.vector(man_north$`Bias Motive Description`) 

bias_man_north<- as.matrix(table(bias_man_north))
bias_man_north

barchart(bias_man_north, main = "Bar Chart of Bias in Manhattan North" , col = "cadetblue2")
```

#### Manhattan South
```{r}
bias_man_south <- as.vector(man_south$`Bias Motive Description`) 

bias_man_south<- as.matrix(table(bias_man_south))
bias_man_south

barchart(bias_man_south, main = "Bar Chart of Bias in Manhattan South" , col = "cadetblue2")
```

#### Queens North
```{r}
bias_queens_north <- as.vector(queens_north$`Bias Motive Description`) 

bias_queens_north<- as.matrix(table(bias_queens_north))
bias_queens_north

barchart(bias_queens_north, main = "Bar Chart of Bias in Queens North" , col = "cadetblue2")
```

#### Queens South
```{r}
bias_queens_south <- as.vector(queens_south$`Bias Motive Description`) 

bias_queens_south<- as.matrix(table(bias_queens_south))
bias_queens_south

barchart(bias_queens_south, main = "Bar Chart of Bias in Queens South" , col = "cadetblue2")
```

#### Staten Island
```{r}
bias_staten_island <- as.vector(staten_island$`Bias Motive Description`) 

bias_staten_island<- as.matrix(table(bias_staten_island))
bias_staten_island

barchart(bias_staten_island, main = "Bar Chart of Bias in Staten Island" , col = "cadetblue2")
```
 
 
  













